---
title: "ASDS5301_FINAL_PROJECT"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
date: "2023-12-06"
---
# TWO WAY ANOVA TEST

## Team 5
## Abhishek Goudar (1001916610)
## JeyaSooriya Sarvanan (1002186838)




```{r}
df_tooth_growth <- data.frame(ToothGrowth)

# Structure of the dataset
str(df_tooth_growth)

head(df_tooth_growth[c(28, 29, 30, 31, 32, 33), ], 
     n = 6)
```
```{r}
##Dataset summary

summary(df_tooth_growth)
```
```{r}

##Descriptive stats in conjunction to both factors
require(dplyr)
library(dplyr)
group_by(df_tooth_growth, supp, dose) %>%
  summarise(
    count = n(),
    mean = mean(len, na.rm = TRUE),
    sd = sd(len, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}

## Boxplot for Delivery Methods of Supplement
par(mar = c(5, 5, 2, 5))
boxplot(len~supp, data=df_tooth_growth, 
        notch=FALSE, 
  col=(c("lightcoral","lightgreen")),
  main="Delivery Methods for Vitamin C Supplements",
  xlab="Supplement Type (Asorbic Acid/Orange Juice)",
  ylab="Length of Odontoblast (Microns)",
  width = c(0.5, 1.5)*2, 
  height = c(0.5, 1.5)
)

```

```{r}
boxplot(len~dose, data=df_tooth_growth, notch=FALSE, 
  col=(c("lightcoral","lightgreen")),
  main="Three Dose Levels of Vitamin C (0.5, 1, and 2)",
  xlab="Numeric Dose in milligrams/day",
  ylab="Length of Odontoblast (Microns)"
  )


```

```{r}
boxplot(len~supp*dose, data=df_tooth_growth, notch=FALSE, 
  col=(c("lightcoral","lightgreen")),
  main="Odontoblast Growth",
  xlab="Suppliment and Dose",
  ylab="Length of Odontoblast (Microns)")
```

```{r}
require(ggplot2)
require(dplyr)
library(ggplot2)
library(dplyr) 

# Convert the 'supp' and 'dose' variables to factors
df_tooth_growth$supp <- as.factor(df_tooth_growth$supp)
df_tooth_growth$dose <- as.factor(df_tooth_growth$dose)


```

```{r}
par(mar = c(8, 8, 2, 8))

# Interaction Plot
interaction.plot(df_tooth_growth$dose, df_tooth_growth$supp, 
                 df_tooth_growth$len, 
                 col = c("lightcoral", "lightgreen"), 
                 type = "b", fixed = TRUE, 
                 xlab = "Dose (milligrams/day)", 
                 ylab = "Length of Odontoblast (Microns)", 
                 pch = 19, leg.bty = "o")

```
```{r}

# Main Effects Plot for 'dose'
ggplot(df_tooth_growth, aes(x = dose, y = len, group = supp, color = supp)) +
  geom_line() +
  geom_point() +
  labs(title = "Main Effects Plot for Dose", 
       x = "Dose (milligrams/day)", y = "Length of Odontoblast (Microns)") +
  scale_color_manual(values = c("lightcoral", "lightgreen")
                     ) +  # Set custom colors
  theme_minimal()

```
```{r}

means_data <- df_tooth_growth %>%
  group_by(supp, dose) %>%
  summarize(mean_len = mean(len),
            .groups="drop")

# Line plot of means
ggplot(means_data, aes(x = dose, y = mean_len, group = supp, color = supp)) +
  geom_line() +
  geom_point(size = 3) +
  labs(title = "Mean Odontoblast Length for Each Supplement vs Dose",
       x = "Dose (milligrams/day)", y = "Mean Odontoblast Length (Microns)") +
  scale_color_manual(values = c("lightcoral", "lightgreen")
                     ) +  # Set custom colors
  theme_minimal()
```


```{r}
# Main Effects Plot for 'supp' 
ggplot(df_tooth_growth, aes(x = supp, y = len, group = dose, color = dose)) +
  geom_boxplot() +
  geom_point(position = position_dodge(width = 0.75), size = 2) +
  labs(title = "Main Effects Plot for Supplement", x = "Supplement", 
       y = "Mean Odontoblast Length (Microns)") +
  scale_color_manual(values = c("lightcoral", "lightgreen", "blue")
                     ) +  # Set custom colors
  theme_minimal()

```


Null Hypothesis: 

Alternate Hypothesis:

Test for Assumptions:

1. Shapiro-Wilk normality test to check for Normality:

Null Hypothesis (H0): The data in the sample is normally distributed.

Alternative Hypothesis (Ha): The data in the sample is not normally distributed.

Used the below code to test for the Normality which is Shapiro.Test.
Taken alpha = 0.05

```{r}
shapiro.test(df_tooth_growth$len)
```

```{r}
qqnorm(df_tooth_growth$len)
qqline(df_tooth_growth$len, col = "red")
```
2.Homogeneity of variance Test
```{r}
require(car)
library(car)
# Levene test for Interaction Term
#levels(ToothGrowth$supp)
# Convert 'dose' to a factor with specific levels
df_tooth_growth$dose <- factor(df_tooth_growth$dose, levels = c("0.5", "1", "2"))

# Perform Levene test
leveneTest(len ~ supp * dose, data = df_tooth_growth)

# Levene test for factor1
leveneTest(len ~ supp, data = df_tooth_growth)

# Levene test for factor2
leveneTest(len ~ dose, data = df_tooth_growth)
```

3. Independence 
```{r}

# Create a contingency table
contingency_table <- table(df_tooth_growth$supp, df_tooth_growth$dose)

# Conduct Chi-squared test for independence
chi_squared_test <- chisq.test(contingency_table)

# Print the results
print(chi_squared_test)

```
Since the Sample contains different categories within the corresponding features, we can confirm that the these features are independent of each other.




```{r}
two_way_anova <- aov(len ~ supp * factor(dose), data = df_tooth_growth)
```

```{r}
summary(two_way_anova)
```

```{r}
TukeyHSD(two_way_anova)
```

```{r}
plot(TukeyHSD(two_way_anova))

```
```{r}
require(rstatix)
library(rstatix)

eta_squared(two_way_anova)


partial_eta_squared(two_way_anova)
```